---
title: "work"
author: "Zhanbo Guo"
date: "`r Sys.Date()`"
output: html_document
---
# Lab 7: ANOVA

## Introduction

This project practises ANOVA models in R. The aim is to compare group means across multiple groups and decide whether there are statistically significant differences.

## Files

- `Computer lab 7.1.R`: original R script for Exercise 1
- `data/`: dataset folder
- `output/`: output figures

## Load data

```{r}
fish <- read.csv("datas/fish1.csv", stringsAsFactors = T)
#stringsAsFactors = TRUE converts text columns into factors, which allows R to treat categorical variables as groups in statistical analyses.
str(fish)   
head(fish)
```
## Boxplot of fish weight by strain

This code creates a boxplot of fish weight for each strain. It allows the distribution, median, spread and possible outliers of weight to be compared across different Tilapia strains.

```{r}
with(fish, boxplot(weight ~ Strain))
```
## Edited boxplot

The boxplot was edited to improve readability. The x-axis and y-axis labels were added, the box width was reduced using `boxwex`, and the y-axis range was set from 0 to 5 using `ylim`.

```{r}
# Create an edited boxplot with axis labels, thinner boxes, and a fixed y-axis range
with(fish, boxplot(weight ~ Strain,
                   xlab = "Tilapia strain",
                   ylab = "Weight",
                   boxwex = 0.5,
                   ylim = c(0, 5)))
```

## Test for equal group variances

Before conducting ANOVA, Bartlett’s test was used to test whether the variance of fish weight was equal across different Tilapia strains.

```{r}
with(fish, bartlett.test(weight ~ Strain))
```
Bartlett’s test showed no significant difference in variance among the three Tilapia strains, χ²(2) = 0.505, p = 0.777. Therefore, we do not reject the null hypothesis. This suggests that the assumption of equal group variance is reasonable, so a standard one-way ANOVA can be used.

## Pairwise t-tests

Because the ANOVA compares all strains together, pairwise t-tests were used to identify which specific Tilapia strains differed from each other. Since Bartlett’s test suggested that the equal variance assumption was reasonable, `pool.sd = TRUE` was used.

```{r}
with(fish, pairwise.t.test(weight, Strain, pool.sd = TRUE))
```

## Result

The pairwise t-test table gives the adjusted p-values for each pairwise comparison between Tilapia strains.

The results suggest that:

- Strain A was significantly different from Strain B, with an adjusted p-value of 0.004.
- Strain A was not significantly different from Strain C, with an adjusted p-value of 0.16.
- Strain B was not significantly different from Strain C, with an adjusted p-value of 0.16.

Therefore, the only statistically significant pairwise difference was between Strain A and Strain B.